Physics-based protein structure prediction and design using the confinement method

### 2012-10-31 Paper/paper.tex
fourth draft

\begin{table}
\label{table:casp_control}

\begin{center}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline 
\end{tabular}
\end{center}

\caption{The confinement method assigns a more favorable free energy to the experimentally
determined structure than to computer-generated predictions. For each target, we examined as many as
five predictions submitted by CASP participants. Positive $\Delta\Delta G$ values indicate that the
experimental structure is predicted to be more favorable than any of the decoys.}

@@

@@

Further Analysis is going on Target 560.

@@

The confinement method has been described in details in ref. by Tyka et.al. and Cecchini et. al.
Here we briefly describe the methodology.  We compute free energy $\Delta G_{AB}$ between A and B
conformation of the protein. For this purpose we use a thermodynamic cycle.

@@

\item  We minimized both the conformation. These minimized conformation(A* and B*) are the
    restrained conformation of that state

\item Backbone dihedral angle of the whole protein is restrained by a harmonic potential to define
    the configurational state.

\item  In order to calculate the free energy $\Delta G_{A,A*}$ and  $\Delta G_{B,B*}$ A and B are
    gradually transformed in A* and B* by applying a harmonic restraint potential to all atom. For
    this purpose around 21 simulation were run each of 20 ns with increasing harmonic restraint
    potential from 0.00005 to 81.92 untill the free and the restrained state overlap well. The final
    restrained state was chosen so high so that the rotational contributional of the protein frozen
    out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
    the reference structure was recorded and the free energy is calculated using a numerical
    approach developed by Tyka et. al.

\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
    restrained state using normal mode analysis.

@@

>>>>

\begin{table}
\label{table:casp_control}
\begin{center}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $12.49 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-2.48 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $22.00 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.24 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $22.15 \pm 0.49$ \\ \hline
     T0569       &    2KWY        &          $-9.40 \pm 0.69$  \\ \hline
\end{tabular}
\end{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally
determined structure than to computer-generated predictions. For each target, we examined as many as
five predictions submitted by CASP participants. Positive $\Delta\Delta G$ values indicate that the
experimental structure is predicted to be more favorable than any of the decoys.}

@@

There is another instance where this method fails. This is the case of solution NMR structure of a domain of adhesion exoprotein 
from Pediococcus pentosaceus. The
pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
best model structure submitted for this target in CASP9. This model structute was submitted by the group " Mufold". ALl the
five model structures submitted by this group was found to be close. Thus we proceed by comparing only the native structure and
the best predicted model of this target. Visualization upon superimposition of two structures reveals that
the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
native structure(see Figure~\ref{fig:T0569}) . Other regions look similar for both the cases with of model1 is within 2.6 \AA
rmsd value of the native structure. In this case we found that the confinement method predict that the predicted model is stable than the
native structure with a free energy difference of $9.4 \pm 0.7$ Kcal/Mol.

@@

\Arijit{Will upload the result from Target 560 once cluster is back.}

@@

The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and 
Cecchini et. al. \cite{Cecchini2009}. The basic approach of confinement approach is same in both
these paper. However, there are some technical differences. Here we briefly describe what we follow
to compute free energy $\Delta G_{AB}$ between A and B
conformation of the protein.

@@

\item  We first minimized both the conformation, A and B. These minimized conformation(A* and B*) 
       are the reference conformation of that state

\item Backbone dihedral angle of both the full protein is restrained by a harmonic potential. This is required to keep
    the configurational state of the protein. This approach is particularly important, so that the any misfolded part 
    of the protein does not chage its configuration.

\item  The free energy is calculated using a thermodynamic cycle. The first part of this approach is the confinement, where
       the state A and B are confined to the reference state A* and B*. This is done gradually by applying larger and larger
       harmonic restraint on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation 
       each of 20 ns long, where the harmonic restraint constant was varied from 0.00005 to 81.92 untill the free and the 
       restrained state overlap well. The final restrained state was chosen so high so that the rotational contributional 
       of the protein frozen out and only remaining contribution remain that of vibrational free enrgy. The fluctuation from
       the reference structure was recorded and the free energy is calculated using a numerical
       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in this way recorder as 
       $\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ as shown in Figure~\ref{fig:method}.

\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
       restrained state A* and B* using normal mode analysis or quasiharmonic analysis. The free energy calculated in 
       this way is shown as $\Delta G_{A*,B*}$ in Figure~\ref{fig:method}.

\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated using the equation 
       $\Delta G_{A,B}$ = $\Delta G_{A,A*}$ - $\Delta G_{B,B*}$ + $\Delta G_{A*,B*}$

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.2 in,height=3.8 in]{T0569.pdf}
\end{center}
\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
The best model was from the group "Mufold".}
\label{fig:T0569}
\end{figure}
